### 2020-03-31 README.Rmd
add quickstart section

R users frequently write scripts which must load data from an external file -- a step which increases friction in reuse and creates a common failure point in reproducibility of the analysis later on. Reading a file directly from a URL is often preferable, since we don't have to worry about distributing the data seperately ourselves.  For example, an analysis might read in the famous CO2 ice core data directly from ORNL repository:

@@

However, we know that data hosted at a given URL could change or dissappear, and not all data we want to work with is available at a URL to begin with.  Digital Object Identifiers (DOIs) were created to deal with these problems of 'link rot'.  Unfortunately, there is no straight forward and general way to read data directly from a DOI, (which almost always resolves to a human-readable webpage rather than the data itself), often apply to collections of files rather than individual source we want to read in our script, and we must frequently work with data that does not (yet) have a DOI. Registering a DOI for a dataset has gotten easier through repositories with simple APIs like Zenodo and figshare, but this is still an involved process and still leaves us without a mechanism to directly access the data.

@@

```{r}
library(contentid)
register("http://cdiac.ornl.gov/ftp/trends/co2/vostok.icecore.co2")
```

@@

We can confirm this is the same data of course:

@@

## Summary

`register()` and `resolve()` provide a low-friction mechanism to create a permanent identifier for external files and then resolve that identifer to an appropriate source.  This can be useful in scripts that are frequently re-run as a way of caching the download step, and simultaneously helps ensure the script is more reproducible.  While this approach is not fail-proof (since all registered locations could fail to produce the content), if all else fails our script itself still contains a cryptographic fingerprint of the data we could use to verify if a given file was really the one used.

## Acknolwedgements

>>>>

## Quick start

Install the current development version using:

```r
# install.packages("remotes")
remotes::install_github("cboettig/contentid")
```
```{r}
library(contentid)
```

Instead of reading in data directly from a local file or URL, use `register()` to register permanent content-based identifiers for your external data file or URL:

```{r}
register("http://cdiac.ornl.gov/ftp/trends/co2/vostok.icecore.co2")
```

Then, `resolve()` that content-based identifier in your scripts for more reproducible workflow. Optionally, set `store=TRUE` to enable local caching:

```{r}
vostok <- resolve("hash://sha256/9412325831dab22aeebdd674b6eb53ba6b7bdd04bb99a4dbb21ddff646287e37",
                  store = TRUE)
```

`resolve` will download and cryptographically verify the identifier matches the content, returning a local file path.  Use that file path in the of our analysis script, e.g.

```{r}
co2 <- read.table(vostok, 
                  col.names = c("depth", "age_ice", "age_air", "co2"), skip = 21)
```

---

## Overview

R users frequently write scripts which must load data from an external file -- a step which increases friction in reuse and creates a common failure point in reproducibility of the analysis later on. Reading a file directly from a URL is often preferable, since we don't have to worry about distributing the data separately ourselves.  For example, an analysis might read in the famous CO2 ice core data directly from ORNL repository:

@@

However, we know that data hosted at a given URL could change or disappear, and not all data we want to work with is available at a URL to begin with.  Digital Object Identifiers (DOIs) were created to deal with these problems of 'link rot'.  Unfortunately, there is no straight forward and general way to read data directly from a DOI, (which almost always resolves to a human-readable webpage rather than the data itself), often apply to collections of files rather than individual source we want to read in our script, and we must frequently work with data that does not (yet) have a DOI. Registering a DOI for a dataset has gotten easier through repositories with simple APIs like Zenodo and figshare, but this is still an involved process and still leaves us without a mechanism to directly access the data.

@@

```{r}
register("http://cdiac.ornl.gov/ftp/trends/co2/vostok.icecore.co2")
```

@@

Note that we have manually embedded the identifier in our script, rather than automatically passing the identifier returned by `register()` directly to resolve.  The command to `register()` needs to only be run once, and thus doesn't need to be embedded in our script (though it is harmless to include it, as it will always return the same identifier unless the data file itself changes).

We can confirm this is the same data:

@@

`register()` and `resolve()` provide a low-friction mechanism to create a permanent identifier for external files and then resolve that identifier to an appropriate source.  This can be useful in scripts that are frequently re-run as a way of caching the download step, and simultaneously helps ensure the script is more reproducible.  While this approach is not fail-proof (since all registered locations could fail to produce the content), if all else fails our script itself still contains a cryptographic fingerprint of the data we could use to verify if a given file was really the one used.

## Acknowledgements
